Boundary terms in the action for Causal Sets

### 2014-10-28 paper_v7.tex
put in sumatis changes to RNC stuff and components of A matrix

Fix $p\in M$ and denote its coordinate values in GNCs by $x^\mu_p= (t_p,\mathbf x_p)$. It will be convenient to use RNCs centered not at the tip, $p$, of the cone but instead at the point $p_0$ where the unique geodesic through $p$ whose tangent is normal to $\Sigma$ intersects $\Sigma$. In GNCs this simply corresponds to the point $p_0$ with coordinates $x_0^\mu=x^\mu (p_0)= (0,\mathbf x_p)$. RNCs centered at $p_0$ will be denoted as $y^{\overline{\mu}}=(y^{\overline{0}},\mathbf{y})=(\overline{t},\mathbf{y})$.
%\mb{let's put all facts about $x^\mu$ and $y^{\overline\mu}$ for $p$ and $p_0$ here and expansion of det in RNC.} 
We need to assume that for any $p$ in $U_\varepsilon$, the Riemann normal neighbourhood $U_p\subset M$ throughout which the RNC system centered at $p_0$ is well-defined contains the truncated cone $J^- (p)\cap J^+ (\Sigma)$. This assumption seems reasonable given the that $U_\varepsilon$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$. The RNCs are constructed such that the metric and the Christoffel symbols at $p_0$ are those of flat space. Following the usual procedure, we define the coordinates $y^{\overline\mu}$ to second order as\mb{sign?}
\be\label{eq:RNCtotaltrans}
y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}(x^\nu -x_0^{\nu})+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)(x^\nu -x_0^{\nu}) (x^\rho -x_0^{\rho})+O ( (x-x_0)^3).
\ee
These coordinates are correctly centred at $p_0$, satisfying $y^{\overline{\mu}}=0$ when $x^{\mu}=x_0^{\mu}$. If the constant matrix $A^{\overline\mu}_{\;\mu}$ obeys 
\be\label{eq:RNCMetricTransAtPAndChris}
A^{\overline\mu}_{\;\mu} A^{\overline\nu}_{\;\nu}\eta_{\overline\mu\overline\nu}=g_{\mu\nu}(p_0)
\ee
the metric and Christoffel symbols take the form
\be\bea
g_{\overline{\mu} \overline{\nu}} (p_0)&=\eta_{\overline\mu\overline\nu}\\\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)&=0.
\eea\ee
The inverse coordinate transformation is
\be\label{eq:RNCinversetrans}
x^{\mu}=x_0^{\mu}+A^{\mu}_{\;\overline{\mu}}y^{\overline{\mu}}+O (y^3)
\ee
where $A^{\mu}_{\;\overline{\mu}}$ is the inverse matrix of $A^{\overline\mu}_{\;{\mu}}$, i.e. $A^{\overline{\mu}}_{\;\mu}A^{\mu}_{\;\overline{\nu}}=\delta^{\overline{\mu}}_{\;\overline{\nu}}$ and $A^{\mu}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\nu}=\delta^{\mu}_{\;\nu}$. There is no $O(y^2)$ term in~\eqref{eq:RNCinversetrans} due to the fact that $\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0$. Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$. Equation~\eqref{eq:RNCMetricTransAtPAndChris} is then satisfied if we set $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $A^{\ibar}_{\;i}A^{\jbar}_{\;j}\delta_{ij} =h_{ij}(p_0)\iff\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$. These relations, along with~\eqref{eq:RNCtotaltrans}, imply that $y^{\overline{0}}_p=\overline{t}_p=T+O(T^3)$.

@@

The base of the cone in GNCs is simply a part of the surface $t=0$, so we can use (\ref{eq:RNCtotaltrans}) to find the equation for the surface in RNCs. Equation (\ref{eq:RNCtotaltrans}) gives
\be\label{eq:BottomSurfaceWithGNC}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)(x^i-x_0^i) (x^j-x_0^j)+O ( (x-x_0)^3)
\ee
The linear part on the right hand side of (\ref{eq:RNCtotaltrans}) vanishes, since we have $A^{\overline{0}}_{\;\mu}(x^{\mu}-x_0^{\mu})=t$ (given that $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $t=0$ on the bottom surface. Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface in RNCs:
\be\label{eq:BottomSurface}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}y^{\ibar} y^{\jbar}+O (y^3)
\ee
Let us rewrite this equation in spherical polar coordinates, i.e. define $r:=\sqrt{\delta_{\ibar\jbar}y^\ibar y^\jbar}$ and the usual angular coordinates $\phi_1,..,\phi_{d-2}$ in terms of the spatial coordinates $y^{\overline{1}} = r \cos (\phi_1),\ldots, y^{\overline{d-1}} = r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})$. Then
\be\label{eq:RadialBottomSurface}
\overline{t}=\frac{1}{2}\left (\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}\frac{y^{\ibar} y^{\jbar}}{r^2}\right)r^2+O (y^3)=\frac{1}{2}f (\mathbf{x}_p,{\boldsymbol\phi})r^2+O (y^3)
\ee
where $\boldsymbol\phi$ stands collectively for all the angular coordinates $\phi_1,..,\phi_{d-2}$. The function $f (\mathbf{x}_p,\boldsymbol\phi)$ depends on $\mathbf{x}_p$ since $\Gamma^{0}_{\;ij}$ and $A^{i}_{\;\ibar}$ depend on $p_0$.
%One can see that $f (\mathbf{x}_0,\boldsymbol\phi)$ depends on the angles, $\phi$, by substituting the relations below into (\ref{eq:RadialBottomSurface})
%\begin{equation}
%\begin{aligned}\label{eq:SphericalCoords}
%y^{\overline{1}} &= r \cos (\phi_1)  \\
%y^{\overline{2}} &= r \sin (\phi_1) \cos (\phi_2)  \\
%y^{\overline{2}} &= r \sin (\phi_1) \sin (\phi_2) \cos (\phi_3)  \\
%    &\vdots  \\
%y^{\overline{d-2}} &= r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \cos (\phi_{d-2})  \\
%y^{\overline{d-1}} &= r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})
%\end{aligned}
%\end{equation}

>>>>

Fix $p\in M$ and denote its coordinate values in GNCs by $x^\mu_p= (t_p,\mathbf x_p)$. It will be convenient to use RNCs centered not at the tip, $p$, of the cone but instead at the point $p_0$ where the unique geodesic through $p$ whose tangent is normal to $\Sigma$ intersects $\Sigma$. In GNCs this simply corresponds to the point $p_0$ with coordinates $x_0^\mu=x^\mu (p_0)= (0,\mathbf x_p)$. RNCs centered at $p_0$ will be denoted as $y^{\overline{\mu}}=(y^{\overline{0}},\mathbf{y})=(\overline{t},\mathbf{y})$.
%\mb{let's put all facts about $x^\mu$ and $y^{\overline\mu}$ for $p$ and $p_0$ here and expansion of det in RNC.} 
We need to assume that for any $p$ in $U_\varepsilon$, the Riemann normal neighbourhood $U_p\subset M$ throughout which the RNC system centered at $p_0$ is well-defined contains the truncated cone $J^- (p)\cap J^+ (\Sigma)$. This assumption seems reasonable given the that $U_\varepsilon$ can be made arbitrarily ``thin'' as $\rho\rightarrow\infty$. The RNCs are constructed such that the metric and the Christoffel symbols at $p_0$ are those of flat space. Following the usual procedure, we define the coordinates $y^{\overline\mu}$ to second order as\mb{sign?}
\be\label{eq:RNCtotaltrans}
y^{\overline{\mu}}=A^{\overline{\mu}}_{\;\nu}(x^\nu -x_0^{\nu})+\frac{1}{2}A^{\overline{\mu}}_{\;\mu}\Gamma^{\mu}_{\;\nu\rho} (p_0)(x^\nu -x_0^{\nu}) (x^\rho -x_0^{\rho})+O ( (x-x_0)^3).
\ee
These coordinates are correctly centred at $p_0$, satisfying $y^{\overline{\mu}}=0$ when $x^{\mu}=x_0^{\mu}$. If the constant matrix $A^{\overline\mu}_{\;\mu}$ obeys 
\be\label{eq:RNCMetricTransAtPAndChris}
A^{\overline\mu}_{\;\mu} A^{\overline\nu}_{\;\nu}\eta_{\overline\mu\overline\nu}=g_{\mu\nu}(p_0)
\ee
the metric and Christoffel symbols take the form
\be\bea
g_{\overline{\mu} \overline{\nu}} (p_0)&=\eta_{\overline\mu\overline\nu}\\\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)&=0.
\eea\ee
The inverse coordinate transformation is
\be\label{eq:RNCinversetrans}
x^{\mu}=x_0^{\mu}+A^{\mu}_{\;\overline{\mu}}y^{\overline{\mu}}+O (y^3)
\ee
where $A^{\mu}_{\;\overline{\mu}}$ is the inverse matrix of $A^{\overline\mu}_{\;{\mu}}$, i.e. $A^{\overline{\mu}}_{\;\mu}A^{\mu}_{\;\overline{\nu}}=\delta^{\overline{\mu}}_{\;\overline{\nu}}$ and $A^{\mu}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\nu}=\delta^{\mu}_{\;\nu}$. There is no $O(y^2)$ term in~\eqref{eq:RNCinversetrans} due to the fact that $\Gamma^{\overline{\mu}}_{\;\overline{\nu}\overline{\rho}} (p_0)=0$. Let us denote the time-coordinate of $p$ in GNCs by $T=x^0_p=t_p$. From \cite{Khetrapal_Sumati:Causal_Diamond_Volume} we have that the tip of the cone in RNC's is $y^{\overline{0}}_p=\overline{t}_p=T+O(T^3)$, which, along with~\eqref{eq:RNCMetricTransAtPAndChris}, sets the components of $A^{\overline{\mu}}_{\mu}$ to be $A^{\overline 0}_{\;0}=1$, $A^{\overline 0}_{\;i}=0$ and $A^{\ibar}_{\;i}A^{\jbar}_{\;j}\delta_{ij} =h_{ij}(p_0)\iff\delta_{\ibar\jbar}=A^i_{\;\ibar}A^j_{\;\jbar}h_{ij} (p_0)$.

@@

The base of the cone in GNCs is simply a part of the surface $t=0$, so we can use (\ref{eq:RNCtotaltrans}) to find the equation for the surface in RNCs. This gives
\be\label{eq:BottomSurfaceWithGNC}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)(x^i-x_0^i) (x^j-x_0^j)+O ( (x-x_0)^3)
\ee
The linear part on the right hand side of (\ref{eq:RNCtotaltrans}) vanishes, since we have $A^{\overline{0}}_{\;\mu}(x^{\mu}-x_0^{\mu})=t$ (given that $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $t=0$ on the bottom surface. Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface in RNCs:
\be\label{eq:BottomSurface}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}y^{\ibar} y^{\jbar}+O (y^3)
\ee
Let us rewrite this equation in spherical polar coordinates, i.e. define $r:=\sqrt{\delta_{\ibar\jbar}y^\ibar y^\jbar}$ and the usual angular coordinates $\phi_1,..,\phi_{d-2}$ in terms of the spatial coordinates $y^{\overline{1}} = r \cos (\phi_1),\ldots, y^{\overline{d-1}} = r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})$. Then
\be\label{eq:RadialBottomSurface}
\overline{t}=\frac{1}{2}\left (\Gamma^{0}_{\;ij} (p_0)A^{i}_{\;\ibar}A^{j}_{\;\jbar}\frac{y^{\ibar} y^{\jbar}}{r^2}\right)r^2+O (y^3)=\frac{1}{2}f (\mathbf{x}_p,{\boldsymbol\phi})r^2+O (y^3)
\ee
where $\boldsymbol\phi$ stands collectively for all the angular coordinates $\phi_1,..,\phi_{d-2}$. The function $f (\mathbf{x}_p,\boldsymbol\phi)$ depends on $\mathbf{x}_p$ since $\Gamma^{0}_{\;ij}$ and $A^{i}_{\;\ibar}$ depend on $p_0$.
%One can see that $f (\mathbf{x}_0,\boldsymbol\phi)$ depends on the angles, $\phi$, by substituting the relations below into (\ref{eq:RadialBottomSurface})
%\begin{equation}
%\begin{aligned}\label{eq:SphericalCoords}
%y^{\overline{1}} &= r \cos (\phi_1)  \\
%y^{\overline{2}} &= r \sin (\phi_1) \cos (\phi_2)  \\
%y^{\overline{2}} &= r \sin (\phi_1) \sin (\phi_2) \cos (\phi_3)  \\
%    &\vdots  \\
%y^{\overline{d-2}} &= r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \cos (\phi_{d-2})  \\
%y^{\overline{d-1}} &= r \sin (\phi_1) \cdots \sin (\phi_{d-3}) \sin (\phi_{d-2})
%\end{aligned}
%\end{equation}
